```{r, include = F}
library(here)
source(file = here::here("functions/functions.R"))
f_load_libraries_colors()
```

# Context of the study {#sec-A}

Environmental risk assessments (ERAs) of chemicals rely on standardized ecotoxicological tests that largely overlook inter-individual variability [@jager_2013b], an essential contributor to ecological resilience and evolutionary success [@desroches_2018; @hendry_2020; @debecker_2019]. Contaminants can disrupt individual differences by altering life-history traits, leading to amplified fitness disparities, favored phenotypes, and a reduced genetic diversity, ultimately affecting population dynamics and its ability to adapt to additional stressors. There is growing consensus that inter-individual variation should be a critical endpoint in ecotoxicological studies, particularly for evaluating long-term impacts on population resilience and community structure [@oziolor_2020]. However, the extent to which pollutants influence inter-individual variability, and how this variability translates into population-level consequences, remains poorly understood.

Behavioral ecotoxicology studies [@polverino_2021; @royaute_2015] suggest two competing scenarios [@killen_2013]: contaminants either constrain individual variation through environmental pressures or amplify it by triggering diverse responses. To address this complexity, researchers now employ sophisticated statistical approaches such as hierarchical models [@feller_2015], enabling better understanding of population responses to pollutants.

The present study focuses on estimating the impact of pesticides exposure on the expression of inter-individual variations in earthworms as key organism for soil functioning [@keith_2012]. Given their ecological niche, earthworms are particularly vulnerable to pesticide residues in soils, and pesticides have been implicated in the decline of earthworm biodiversity in conventionally managed agricultural fields [@bertrand_2024]. We specifically focus on *Aporrectodea caliginosa*, a model species used in ecotoxicology; which is ubiquist in agricultural fields [@bart_2018b]. *A. caliginosa* is also one of the model species for ISO ecotoxicological tests (ISO 11268-2:2023).

Our study had two primary objectives :

-   To investigate how pesticide exposure affects not only the mean growth rate but also the inter-individual variations in earthworm growth, with important implications for environmental risk assessment (ERA).
-   To evaluate whether current ERA protocols, which typically use fewer replicates, can reliably capture these biological variations.

To address these objectives, we conducted a longitudinal study with two treatments: one control and one with a sub-lethal dose of a cocktail of epoxiconazole and dimoxystrobin, two active substances found in the plant protection product Swing® Gold. This fungicide has known effects on A. caliginosa growth and reproduction [@bart_2019c; @bart_2020a].

Because of natural phenotypic diversity, individual variation is already expected in the non exposed cohort. We expected a decrease in the average individual growth rate in the fungicide cocktail. We also hypothesized an effect of pesticides exposure on individual variation but with no a priori on the direction of the pesticide effect, as empirical results cited above support both cases (@fig-hyp). To evaluate current ERA practices, we used a resampling procedure to examine whether a reduced sample size would have altered our estimates, potentially leading to inaccurate or even misleading conclusions. We anticipated that standard replicate numbers may not be sufficient to accurately capture both variability and mean growth, increasing the risk of overlooking a particularly sensitive subset of the population.

Through the longitudinal comparison of 30 exposed and 30 unexposed individuals, we discovered that while the fungicide had minimal impact on mean growth rate, it caused a three-fold increase in inter-individual variability, revealing a subset of highly sensitive individuals with up to 10% growth reduction. Our findings have significant implications for environmental risk assessment. Focusing solely on population means could overlook effects on sensitive individuals that may signal early environmental stress. While incorporating individual variability into ecotoxicological studies remains challenging, these efforts are crucial for refining safety factors in ERA frameworks [@jager_2013b; @nikinmaa_2019] and improving risk assessment accuracy. Additionally, better understanding of the impact of contaminants on inter-individual variation would improve the accuracy and ecological relevance of risk assessments, while capturing the long-term implications for population and ecosystem dynamics.

```{r, message=F, warning=F}

col_ctrl <- pal_col[1]
col_trt <- pal_col[2]

Exposition <- c("Control", "Treatment")

size_line <- 1.5
alpha_line <- 0.5
size_text <- 5
size_point <- 5
alpha_point <- 1
size_title <- 13
mar <- 0.2
col_grey <- "grey50"

y_min <- -1
y_max <- 0.5

y_label <- -1.2

mu_dim <- 0.4


plot_Inf <- ggplot()+
  geom_point(
    aes(
    x=Exposition[1], 
    y=c(0)
    ), 
    color=col_ctrl,
    size=size_point, 
    alpha=alpha_point
    )+
  geom_point(
    aes(
      x=Exposition[2], 
      y=c(-mu_dim)), 
    color=col_trt,
    size=size_point, 
    alpha=alpha_point
    )+
  ylim(
    y_min,
    y_max
    )+
  labs(y="Growth rate")+
  annotate(
    "segment",
    x = Exposition[1], 
    xend = Exposition[1],
    y = -0.3,
    yend = 0.3,
    colour = col_ctrl, 
    size=size_line, 
    alpha=alpha_line
    )+
  annotate(
    "segment",
    x = Exposition[2], 
    xend = Exposition[2],
    y = -0.15-mu_dim, 
    yend = 0.15-mu_dim,
    colour = col_trt, 
    size=size_line, 
    alpha=alpha_line
    )+
  labs(subtitle = "1. Constraints on individual \ndifferences in growth")+
  theme_classic(14)+
  theme(
    plot.title = element_text(
      face="bold", 
      size=size_title,
      hjust=0.5, 
      vjust=0.5
      ),
    plot.subtitle = element_text(
      face="plain", 
      size=size_title-2,
      hjust=0.5, 
      vjust=0.5
      ),
    plot.margin = unit(
      c(mar, mar, mar, mar),
      "inches"
      ),
    axis.ticks.y = element_blank(),
    axis.text.y = element_blank(),
    #axis.line.x = element_blank(),
    axis.line.y = element_line(
      arrow = grid::arrow(
        length = unit(0.3, "cm"),
        ends = "last"
        )
      ),
    axis.title.x = element_blank()
    )

plot_Sup <- ggplot()+
  geom_point(
    aes(
      x=Exposition[1], 
      y=c(0)
      ), 
    color=col_ctrl,
    size=size_point, 
    alpha=alpha_point
    )+
  geom_point(
    aes(
      x=Exposition[2], 
      y=c(-mu_dim)
      ), 
    color=col_trt, 
    size=size_point, 
    alpha=alpha_point
    )+
  ylim(
    y_min,
    y_max
    )+
  annotate(
    "segment", 
    x = Exposition[1], 
    xend = Exposition[1],
    y = -0.3,
    yend = 0.3, 
    colour = col_ctrl, 
    size=size_line, 
    alpha=alpha_line
    )+
  annotate(
    "segment",
    x = Exposition[2], 
    xend = Exposition[2],
    y = -0.6-mu_dim, 
    yend = 0.6-mu_dim,
    colour = col_trt, 
    size=size_line, 
    alpha=alpha_line
    )+

  labs(subtitle = "2. Amplification of individual \ndifferences in growth")+
  theme_classic(14)+
  theme(
    plot.title = element_text(
      face="bold", 
      size=size_title,
      hjust=0.5, 
      vjust=0.5
      ),
    plot.subtitle = element_text(
      face="plain", 
      size=size_title-2,
      hjust=0.5, vjust=0.5
      ),
    plot.margin = unit(
      c(mar, mar, mar, mar),
      "inches"
      ),
    axis.ticks.y = element_blank(),
    axis.text.y = element_blank(),
    axis.line.y = element_blank(),
    axis.title = element_blank()
    )

# Population simulation for a decrease in variance
nb_ind <- 10
t_expe <- 56

t_values = seq(0, t_expe, 1) # Time points
sd = 0.000001 # Measurement error

# Biological parameters (Bart et al. 2020)
L0_mu = 2.12                # Initial volumetric structural length (mg1/3) (= 2.25)
a_ctrl_mu = 0.085           # Growth rate (d-1)
a_trt_mu = 0.085*0.7        # Growth rate (d-1)

L0_sd = 0.26
a_ctrl_sd = 0.01
a_trt_sd = 0.005

set.seed(3333)

ID_i  <- seq(1,nb_ind,1)
List_L0_i<- rnorm(nb_ind, L0_mu, L0_sd)
List_a_i <- c(rnorm(nb_ind/2, a_ctrl_mu, a_ctrl_sd), rnorm(nb_ind/2, a_trt_mu, a_trt_sd))

df_pop_sim <- data.frame(                                
  ID = ID_i,                           
  L0_i = List_L0_i,                                        
  a_i = List_a_i,                     
  exposition = as.factor(c(                                             
    rep(0, nb_ind/2),                                       
    rep(1, nb_ind/2)                                        
    ))                                                       
  ) 

df_sim <- df_pop_sim |> 
  expand_grid(t = t_values) |>   
  mutate(
    L = L0_i + a_i * t
    )
data_sim <- df_sim |> 
  mutate(
    w = L^3,
    exposition = as.factor(exposition)
  )

df_sim_decrease <- data_sim

# Population simulation for an increase in variance

a_ctrl_sd = 0.01
a_trt_sd = 0.025

set.seed(3333)

ID_i  <- seq(1,nb_ind,1)
List_L0_i<- rnorm(nb_ind, L0_mu, L0_sd)
List_a_i <- c(rnorm(nb_ind/2, a_ctrl_mu, a_ctrl_sd), rnorm(nb_ind/2, a_trt_mu, a_trt_sd))

df_pop_sim <- data.frame(                                
  ID = ID_i,                           
  L0_i = List_L0_i,                                        
  a_i = List_a_i,                     
  exposition = as.factor(c(                                             
    rep(0, nb_ind/2),                                       
    rep(1, nb_ind/2)                                        
    ))                                                       
  )  

df_sim <- df_pop_sim |> 
  expand_grid(t = t_values) |>   
  mutate(
    L = L0_i + a_i * t
    )
data_sim <- df_sim |> 
  mutate(
    w = L^3,
    exposition = as.factor(exposition)
  )

df_sim_increase <- rbind(subset(df_sim_decrease, exposition==0),subset(data_sim, exposition==1))

alpha_line <- 0.8
line_width <- 1.2

plot_Inf_traj <- ggplot(
  data=df_sim_decrease, 
  aes(
    x=t, 
    y=L^3, 
    color=exposition, 
    group=as.factor(ID))
  )+
  scale_color_manual(values=pal_col, label=Exposition, name="")+
  geom_line(
    lwd=line_width, 
    alpha=alpha_line
    )+
  labs(
    x="Time", 
    y="Weight",
    title = "Consequences on growth",
    subtitle = "1. Low to no overlap in \ngrowth trajectories"
    )+
  theme_classic(14)+
  theme(
    plot.title = element_text(
      face="bold", 
      size=size_title,
      vjust=0.5
      ),
    plot.subtitle = element_text(
      face="plain", 
      size=size_title-2,
      vjust=0.5
      ),
    plot.margin = unit(
      c(mar, mar, mar, mar),
      "inches"
      ),
    axis.ticks = element_blank(),
    legend.position = "bottom",
    axis.text = element_blank(),
    axis.title = element_text(),
    axis.line.x=element_line(),
    axis.line.y = element_line(
      arrow = grid::arrow(
        length = unit(0.3, "cm"),
        ends = "last"
        )
      )
    )

plot_Sup_traj <- ggplot(
  data=df_sim_increase, 
  aes(
    x=t, 
    y=L^3, 
    color=exposition, 
    group=as.factor(ID))
  )+
  scale_color_manual(values=pal_col, label=Exposition, name="")+

  geom_line(
    lwd=line_width, 
    alpha=alpha_line
    )+
  
  labs(
    x="Time", 
    y="",
    subtitle = "2. Strong overlap in growth trajectories \ndespite average differences"
    )+
  theme_classic(14)+
  theme(
    plot.title = element_text(
      face="bold", 
      size=size_title,
      hjust=0.5, 
      vjust=0.5
      ),
    plot.subtitle = element_text(
      face="plain", 
      size=size_title-2,
      vjust=0.5
      ),
    plot.margin = unit(
      c(mar, mar, mar, mar),
      "inches"
      ),
    axis.ticks = element_blank(),
    legend.position = "bottom",
    axis.text = element_blank(),
    axis.title = element_text(),
    axis.line.y = element_blank(),
    axis.line.x = element_line(
      arrow = grid::arrow(
        length = unit(0.3, "cm"),
        ends = "last"
        )
      )
    )
```

```{r, message=F, warning=F}
#| fig-cap: Different scenarios of how pesticides exposure may alter individual differences in growth under the general hypothesis that average growth is reduced by the exposition. Case 1 describes a diminution of individual differences that may be due to the additional constraints from the exposition, leading to a low overlap in growth trajectories between the two groups. Case 2 describes an amplification of individual differences that may be due to the additional variation of individual responses to the fungicide exposure, leading to a strong overlap in growth trajectories between the two groups.
#| label: fig-hyp
#| fig-height: 8
#| fig-width: 8

plot_hyp <- plot_Inf+plot_Sup+ plot_Inf_traj+plot_Sup_traj+plot_layout(ncol=2, guides="collect")+
  plot_annotation(title = "Growth rate comparison") &
  theme(legend.position = "bottom",
        plot.title = element_text(face="bold", size=size_title,
                                  vjust=0.5))
plot_hyp
```

```{r, include=F}
ggsave(
  plot_hyp, 
  filename="plot_hyp.png", 
  path=here::here("fig/"), 
  width=7, 
  height=9
  )
```
