```{r, include = F}
library(tidyverse); library(here); library(ggthemes)
library(viridis); library(tidybayes); library(wesanderson)
library(ggdist); library(patchwork); library(extrafont)
library(scales); library(ggdist); library(ggsci)
library(parallel); library(deSolve); library(tmvtnorm)
library(fdrtool); library(readxl); library(here); library(bayestestR)
library(RColorBrewer); library(devtools); library(car); library(plotly)
library(knitr); library(Hmisc); library(nlstools); library(lsmeans)

col_blue <- "#5E81AC"
col_red <- "#f42404"
pal_col <- c(col_blue, col_red)

getwd()
```

# Context of the study

## Abstract

Environmental risk assessments (ERAs) of chemicals rely on standardized ecotoxicological tests that largely overlook inter-individual variability, an essential contributor to ecological resilience and evolutionary success. Contaminants can disrupt individual differences by altering life-history traits, leading to amplified fitness disparities, favored phenotypes, and a reduced genetic diversity, ultimately affecting population dynamics and its ability to adapt to additional stressors. However, the extent to which pollutants influence inter-individual variability, and how this variability translates into population-level consequences, remains poorly understood.

Addressing this gap, our study assesses the inter-individual variability of growth response in the earthworm Aporrectodea caliginosa to a sublethal exposure to Swing® Gold fungicide, and explores its consequences on individual growth trajectories. Using a longitudinal study design of 30 exposed and 30 unexposed individuals, we compared mean growth rates and inter-individual variability between the two groups. While the fungicide had a weak effect on the cohort’s mean growth rate, we noticed a three-fold increase in inter-individual variability. This increase revealed a subset of highly sensitive individuals with growth reduced up to 10% compared to the average response.

Our results indicate that focusing on the population mean could overlook effects on the most sensitive individuals, while these may signal early environmental stress. Incorporating individual variability into ecotoxicological studies remains challenging due to the labor-intensive nature of individual monitoring, advanced statistical requirements, and the need for larger datasets. Nevertheless, these efforts are crucial, as these studies could be used into higher-tier ERA framework to refine the safety factors accounting for intraspecies variability when defining regulatory thresholds. Additionally, better understanding of the impact of contaminants on inter-individual variation would improve the accuracy and ecological relevance of risk assessments, while capturing the long-term implications for population and ecosystem dynamics.

## Introduction

Environmental risk assessment of chemicals is based on standardized ecotoxicological tests on various model species. These tests were designed under the strong assumption that species sensitivity to contaminants is well represented by the sensitivity of individuals on average, with ecotoxicological parameters being defined at the species level. However, inter-individual variations in life history traits, behavior and physiology within a single population has been repeatedly demonstrated by ecologists and evolutionary biologists and can be as strong as among-species differences according to a recent meta-analysis [@desroches_2018].

The diversity in phenotypic expressions within populations arises from two primary processes: phenotypic plasticity and adaptation [@hendry_2020]. Phenotypic plasticity enables individuals to adjust their phenotype in response to environmental changes during their lifetime. These modifications allow for rapid and flexible responses to dynamic conditions, enhancing an individual’s immediate survival and performance. In contrast, adaptation results from genetic changes that confer a fitness advantage, allowing beneficial phenotypic traits to become established and passed down across generations. Although this process is slower, it ensures the long-term incorporation of advantageous traits within a population. Together, these mechanisms contribute to a broad spectrum of individual responses to environmental changes, shaping both short-term resilience and long-term evolutionary potential [@debecker_2019]. Experimental studies reviewed by @forsman_2016 demonstrated that such variations reduce populations’ fluctuations and vulnerability to environmental changes, enhance colonization success, expand geographic range , and decrease the risk of extinction.

Because the environment partially drives inter-individual differences, anthropogenic stressors, such as pollutants, can impose selective pressures on chemically exposed populations, often leading to accelerated phenotypic changes [@sanderson_2022a]. Pollutants disrupt normal physiological and developmental processes, altering life-history traits and behaviors with cascading effects on population dynamics and ecosystem health [@brown_2009; @relyea_2006]. Although adaptation to pollutants can offer short-term survival benefits, it often entails costs, including reduced genetic variability and fitness trade-offs that impair tolerance to secondary stressors [@coutellec_2011].

Classical ecotoxicology studies have traditionally focused on species-level responses, treating inter-individual variability as noise. Most approaches rely heavily on standardized tests with model organisms and small sample sizes, further restricting the estimation of intra-population variability [@jager_2013b]. This is also evident in biomarker studies, where low-variability indicators are often selected to enhance contaminant effect detection [@depledge_1994; @jemec_2010]. While these approaches facilitate reproducibility, they overlook the range of potential responses within populations and hinder a deeper understanding of individual differences. Even when variations in sensitivity are documented—such as differences in maximum size reductions in Folsomia candida exposed to cadmium [@jager_2013b]—they are rarely incorporated as endpoints in regulatory risk assessments, reducing their relevance for evaluating ecological impacts.

In contrast, behavioral ecotoxicology has made significant strides in understanding individual variability, revealing that contaminants like pesticides and anxiolytics can alter behaviors across diverse taxa, from spiders to odonates and fish [@polverino_2021; @royaute_2015]. However, the magnitude and direction of these effects vary drastically from study to study, emphasizing the importance of accounting for personality traits [@reale_2007] and their disruption by pollutants. Two competing scenarios emerge from theoretical frameworks: exposure to contaminants could either constrain individual variation through additional environmental pressures, or amplify it by triggering diverse individual responses to the exposure (Figure 1) [@killen_2013]. To address this complexity, the field has adopted sophisticated statistical approaches, particularly hierarchical models and reaction norm-based analyses. These statistical tools, by incorporating individual-level data and estimating inter-individual variations through hyperparameters [@feller_2015], enable a more nuanced understanding of how populations respond to pollutants, accounting for the full spectrum of individual variations in both exposure and response.

Consequently, environmental risk assessments (ERAs) based on average responses at population level, may underestimate the actual impact of contaminants, particularly when subpopulations of more sensitive individuals are present. While practical, this approach limits the understanding of individual differences and their ecological significance. There is a growing consensus among researchers that inter-individual variation should be incorporated as a critical endpoint in ecotoxicological studies. Acknowledging and quantifying these variations can improve the predictive power of ecotoxicological models and provide a more accurate assessment of the ecological risks associated with contaminant exposure [@jager_2013b; @nikinmaa_2019]. By incorporating individual variability into population dynamics models, researchers can account for the diverse range of responses to environmental stressors within a population. This approach could be particularly valuable for long-term risk assessments that consider chronic exposures and sub-lethal effects, as variability in traits such as growth, reproduction, and survival can influence the resilience or vulnerability of a population, leading to shifts in community structure and function. The integration of this variability into models would provide a more accurate depiction of population-level outcomes, helping to forecast potential ecological consequences of contaminant exposure more reliably [@oziolor_2020].

The present study focuses on estimating the impact of pesticides exposure on the expression of inter-individual variations in earthworms. Earthworms are key organisms for soil functioning, contributing to air and water flow regulation, soil structuration, and organic matter decomposition [@keith_2012]. Given their ecological niche, earthworms are particularly vulnerable to pesticide residues in soils, and pesticides have been implicated in the decline of earthworm biodiversity in conventionally managed agricultural fields [@bertrand_2024]. We specifically focus on Aporrectodea caliginosa, a model species used in ecotoxicology; which is ubiquist in agricultural fields [@bart_2018b]. A. caliginosa is also one of the model species for ISO ecotoxicological tests (ISO 11268-2:2023).

Our study had two primary objectives. First, we aimed to investigate how pesticide exposure affects not only the mean growth rate but also the inter-individual variations in earthworm growth, with important implications for environmental risk assessment (ERA). Second, we sought to evaluate whether current ERA protocols, which typically use fewer replicates, can reliably capture these biological variations.

To address these objectives, we conducted a longitudinal study with two treatments: one control and one with a sub-lethal dose of a cocktail of epoxiconazole and dimoxystrobin, two active substances found in the plant protection product Swing® Gold. This fungicide has known effects on A. caliginosa growth and reproduction [@bart_2019c; @bart_2020a]. Because of natural phenotypic diversity, individual variation is already expected in the non exposed cohort. We expected a decrease in the average individual growth rate in the fungicide cocktail. We also hypothesized an effect of pesticides exposure on individual variation but with no a priori on the direction of the pesticide effect, as empirical results cited above support both cases (@fig-hyp). To evaluate current ERA practices, we used a resampling procedure to examine whether a reduced sample size would have altered our estimates, potentially leading to inaccurate or even misleading conclusions. We anticipated that standard replicate numbers may not be sufficient to accurately capture both variability and mean growth, increasing the risk of overlooking a particularly sensitive subset of the population.

```{r, message=F, warning=F}

col_ctrl <- pal_col[1]
col_trt <- pal_col[2]

Exposition <- c("Control", "Treatment")

size_line <- 1.5
alpha_line <- 0.5
size_text <- 5
size_point <- 5
alpha_point <- 1
size_title <- 13
mar <- 0.2
col_grey <- "grey50"

y_min <- -1
y_max <- 0.5

y_label <- -1.2

mu_dim <- 0.4


plot_Inf <- ggplot()+
  geom_point(
    aes(
    x=Exposition[1], 
    y=c(0)
    ), 
    color=col_ctrl,
    size=size_point, 
    alpha=alpha_point
    )+
  geom_point(
    aes(
      x=Exposition[2], 
      y=c(-mu_dim)), 
    color=col_trt,
    size=size_point, 
    alpha=alpha_point
    )+
  ylim(
    y_min,
    y_max
    )+
  labs(y="Growth rate")+
  annotate(
    "segment",
    x = Exposition[1], 
    xend = Exposition[1],
    y = -0.3,
    yend = 0.3,
    colour = col_ctrl, 
    size=size_line, 
    alpha=alpha_line
    )+
  annotate(
    "segment",
    x = Exposition[2], 
    xend = Exposition[2],
    y = -0.15-mu_dim, 
    yend = 0.15-mu_dim,
    colour = col_trt, 
    size=size_line, 
    alpha=alpha_line
    )+
  labs(subtitle = "1. Constraints on individual \ndifferences in growth")+
  theme_classic(14)+
  theme(
    plot.title = element_text(
      face="bold", 
      size=size_title,
      hjust=0.5, 
      vjust=0.5
      ),
    plot.subtitle = element_text(
      face="plain", 
      size=size_title-2,
      hjust=0.5, 
      vjust=0.5
      ),
    plot.margin = unit(
      c(mar, mar, mar, mar),
      "inches"
      ),
    axis.ticks.y = element_blank(),
    axis.text.y = element_blank(),
    #axis.line.x = element_blank(),
    axis.line.y = element_line(
      arrow = grid::arrow(
        length = unit(0.3, "cm"),
        ends = "last"
        )
      ),
    axis.title.x = element_blank()
    )

plot_Sup <- ggplot()+
  geom_point(
    aes(
      x=Exposition[1], 
      y=c(0)
      ), 
    color=col_ctrl,
    size=size_point, 
    alpha=alpha_point
    )+
  geom_point(
    aes(
      x=Exposition[2], 
      y=c(-mu_dim)
      ), 
    color=col_trt, 
    size=size_point, 
    alpha=alpha_point
    )+
  ylim(
    y_min,
    y_max
    )+
  annotate(
    "segment", 
    x = Exposition[1], 
    xend = Exposition[1],
    y = -0.3,
    yend = 0.3, 
    colour = col_ctrl, 
    size=size_line, 
    alpha=alpha_line
    )+
  annotate(
    "segment",
    x = Exposition[2], 
    xend = Exposition[2],
    y = -0.6-mu_dim, 
    yend = 0.6-mu_dim,
    colour = col_trt, 
    size=size_line, 
    alpha=alpha_line
    )+

  labs(subtitle = "2. Amplification of individual \ndifferences in growth")+
  theme_classic(14)+
  theme(
    plot.title = element_text(
      face="bold", 
      size=size_title,
      hjust=0.5, 
      vjust=0.5
      ),
    plot.subtitle = element_text(
      face="plain", 
      size=size_title-2,
      hjust=0.5, vjust=0.5
      ),
    plot.margin = unit(
      c(mar, mar, mar, mar),
      "inches"
      ),
    axis.ticks.y = element_blank(),
    axis.text.y = element_blank(),
    axis.line.y = element_blank(),
    axis.title = element_blank()
    )

# Population simulation for a decrease in variance
nb_ind <- 10
t_expe <- 56

t_values = seq(0, t_expe, 1) # Time points
sd = 0.000001 # Measurement error

# Biological parameters (Bart et al. 2020)
L0_mu = 2.12                # Initial volumetric structural length (mg1/3) (= 2.25)
a_ctrl_mu = 0.085           # Growth rate (d-1)
a_trt_mu = 0.085*0.7        # Growth rate (d-1)

L0_sd = 0.26
a_ctrl_sd = 0.01
a_trt_sd = 0.005

set.seed(3333)

ID_i  <- seq(1,nb_ind,1)
List_L0_i<- rnorm(nb_ind, L0_mu, L0_sd)
List_a_i <- c(rnorm(nb_ind/2, a_ctrl_mu, a_ctrl_sd), rnorm(nb_ind/2, a_trt_mu, a_trt_sd))

df_pop_sim <- data.frame(                                
  ID = ID_i,                           
  L0_i = List_L0_i,                                        
  a_i = List_a_i,                     
  exposition = as.factor(c(                                             
    rep(0, nb_ind/2),                                       
    rep(1, nb_ind/2)                                        
    ))                                                       
  ) 

df_sim <- df_pop_sim |> 
  expand_grid(t = t_values) |>   
  mutate(
    L = L0_i + a_i * t
    )
data_sim <- df_sim |> 
  mutate(
    w = L^3,
    exposition = as.factor(exposition)
  )

df_sim_decrease <- data_sim

# Population simulation for an increase in variance

a_ctrl_sd = 0.01
a_trt_sd = 0.025

set.seed(3333)

ID_i  <- seq(1,nb_ind,1)
List_L0_i<- rnorm(nb_ind, L0_mu, L0_sd)
List_a_i <- c(rnorm(nb_ind/2, a_ctrl_mu, a_ctrl_sd), rnorm(nb_ind/2, a_trt_mu, a_trt_sd))

df_pop_sim <- data.frame(                                
  ID = ID_i,                           
  L0_i = List_L0_i,                                        
  a_i = List_a_i,                     
  exposition = as.factor(c(                                             
    rep(0, nb_ind/2),                                       
    rep(1, nb_ind/2)                                        
    ))                                                       
  )  

df_sim <- df_pop_sim |> 
  expand_grid(t = t_values) |>   
  mutate(
    L = L0_i + a_i * t
    )
data_sim <- df_sim |> 
  mutate(
    w = L^3,
    exposition = as.factor(exposition)
  )

df_sim_increase <- rbind(subset(df_sim_decrease, exposition==0),subset(data_sim, exposition==1))

alpha_line <- 0.8
line_width <- 1.2

plot_Inf_traj <- ggplot(
  data=df_sim_decrease, 
  aes(
    x=t, 
    y=L^3, 
    color=exposition, 
    group=as.factor(ID))
  )+
  scale_color_manual(values=pal_col, label=Exposition, name="")+
  geom_line(
    lwd=line_width, 
    alpha=alpha_line
    )+
  labs(
    x="Time", 
    y="Weight",
    title = "Consequences on growth",
    subtitle = "1. Low to no overlap in \ngrowth trajectories"
    )+
  theme_classic(14)+
  theme(
    plot.title = element_text(
      face="bold", 
      size=size_title,
      vjust=0.5
      ),
    plot.subtitle = element_text(
      face="plain", 
      size=size_title-2,
      vjust=0.5
      ),
    plot.margin = unit(
      c(mar, mar, mar, mar),
      "inches"
      ),
    axis.ticks = element_blank(),
    legend.position = "bottom",
    axis.text = element_blank(),
    axis.title = element_text(),
    axis.line.x=element_line(),
    axis.line.y = element_line(
      arrow = grid::arrow(
        length = unit(0.3, "cm"),
        ends = "last"
        )
      )
    )

plot_Sup_traj <- ggplot(
  data=df_sim_increase, 
  aes(
    x=t, 
    y=L^3, 
    color=exposition, 
    group=as.factor(ID))
  )+
  scale_color_manual(values=pal_col, label=Exposition, name="")+

  geom_line(
    lwd=line_width, 
    alpha=alpha_line
    )+
  
  labs(
    x="Time", 
    y="",
    subtitle = "2. Strong overlap in growth trajectories \ndespite average differences"
    )+
  theme_classic(14)+
  theme(
    plot.title = element_text(
      face="bold", 
      size=size_title,
      hjust=0.5, 
      vjust=0.5
      ),
    plot.subtitle = element_text(
      face="plain", 
      size=size_title-2,
      vjust=0.5
      ),
    plot.margin = unit(
      c(mar, mar, mar, mar),
      "inches"
      ),
    axis.ticks = element_blank(),
    legend.position = "bottom",
    axis.text = element_blank(),
    axis.title = element_text(),
    axis.line.y = element_blank(),
    axis.line.x = element_line(
      arrow = grid::arrow(
        length = unit(0.3, "cm"),
        ends = "last"
        )
      )
    )
```

```{r, message=F, warning=F}
#| fig-cap: Different scenarios of how pesticides exposure may alter individual differences in growth under the general hypothesis that average growth is reduced by the exposition. Case 1 describes a diminution of individual differences that may be due to the additional constraints from the exposition, leading to a low overlap in growth trajectories between the two groups. Case 2 describes an amplification of individual differences that may be due to the additional variation of individual responses to the fungicide exposure, leading to a strong overlap in growth trajectories between the two groups.
#| label: fig-hyp
#| fig-height: 8
#| fig-width: 8

plot_hyp <- plot_Inf+plot_Sup+ plot_Inf_traj+plot_Sup_traj+plot_layout(ncol=2, guides="collect")+
  plot_annotation(title = "Growth rate comparison") &
  theme(legend.position = "bottom",
        plot.title = element_text(face="bold", size=size_title,
                                  vjust=0.5))
plot_hyp
```

```{r, include=F}
ggsave(
  plot_hyp, 
  filename="plot_hyp.png", 
  path=here::here("fig/"), 
  width=7, 
  height=9
  )
```
